---
title: "Hands-on"
subtitle: "Advanced data manipulation"
author: "Larisa M. Soto"
date: "2022-09-14"
output:
   rmdformats::html_clean:
    toc: false
    thumbnails: false
    highlight: kate
    use_bookdown: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=TRUE)
```

```{r,echo=FALSE}
htmltools::img(src = knitr::image_uri("Micm_color_logo.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:-10px; right:0; padding:10px; max-width:50%;')
```

```{r}
library(gapminder)
library(plyr)
library(dplyr)
```


# Split-Apply-Combine

Calculate the average life expectancy per continent using `ddply`. Write the result into a new data frame and rename the column with the mean life expectancy with something meaningful. 

```{r}
plyr::ddply(.data = gapminder,
            .variables = c("continent"),
            .fun = function(x) mean(x$lifeExp)
            )
```
Calculate the average life expectancy per continent per year  using `ddply`. Write the result into a new data frame and rename the column with the mean life expectancy with something meaningful. 

```{r}
mle<-plyr::ddply(.data = gapminder,
                 .variables = c("year","continent"),
                 .fun = function(df) mean(df$lifeExp))
colnames(mle)[colnames(mle)=="V1"]<-"meanLifeExp"
head(mle)
```

# Manipulation with `dplyr`

Write one command ( can span multiple lines) using pipes that will output a data frame that has only the columns `lifeExp`, `country` and `year` for the records before the year 2000 from African countries, but not for other Continents.

```{r}
tidy_africa <- gapminder %>%
                dplyr::filter(continent == "Africa") %>%
                dplyr::select(year, country, lifeExp)
head(tidy_africa)
```

Calculate the average life expectancy per country. Which country has the longest average life expectancy and which one the shortest average life expectancy?

```{r}
gapminder %>%
   dplyr::group_by(country) %>%
   dplyr::summarize(mean_lifeExp = mean(lifeExp)) %>%
   dplyr::filter(mean_lifeExp == min(mean_lifeExp) | mean_lifeExp == max(mean_lifeExp))
```

In the previous hands-on you discovered that all the entries from 2007 are actually from 2008. Write a command to edit the data accordingly using pipes. In the same command filter only the entries from 2008 to verify the change.

```{r}
gapminder %>%
  dplyr::mutate(year = ifelse(year==2007,2008,year)) %>%
  dplyr::filter(year==2008) %>%
  head()
```

